### 2024-05-22 duality-presentation.qmd

---
title: "Coasting on Duality"
subtitle: "What We Get When We Do Two-Mode Network CA"
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
slide-number: true
slide-level: 2
progress: true
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    footer: "https://github.com/olizardo/correspondence-analysis-two-mode-networks"
---

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership

::: aside
*Breiger 2000; Roberts 2000; Faust 2005; D’Esposito et al. 2014

**Bonacich 1991; Borgatti & Everett 1997
:::

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership
- CA Proponents point to some important advantages of CA for the analysis of two-mode networks:
  - CA can provide a "joint display" of the structural connectivity patterns in two-mode data
  - (M)CA can be used to "visually explore" two-mode networks

::: aside
*Breiger 2000; Roberts 2000; Faust 2005; D’Esposito et al. 2014

**Bonacich 1991; Borgatti & Everett 1997
:::

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership
- This presentation builds on and expands on previous  research on using CA for two-mode data analysis but discusses aspects of CA that have not been emphasized in previous research:
  - CA as computing a type of "(Reflective) Dual Centrality" on the entities in the two-modes*

:::aside
*Rediscovered recently under the heading of the Economic Complexity Index (ECI; Hidalgo and Hausman 2009)
:::

## CA and Two Mode Networks

- While still not very common, Correspondence Analysis (CA) has been sporadically used in *analyzing* two-mode network data*
- But criticized for being limited and inapplicable for most "SNA" tasks**
  - Computing (Centrality-Like) Measures of Node Position
  - Computing Measures of Group Membership
- This presentation builds on and expands on previous  research on using CA for two-mode data analysis but discusses aspects of CA that have not been emphasized in previous research:
  - CA as extracting "positional" information the entities in the two modes
    - With positions defined by "Generalized Relational Similarity"
    - Similarly central entities have similar patterns of connectivity with entities that are themselves similar**

:::aside
**Jeh & Widom 2002; Kovacs 2010; Lizardo 2024
:::

## Southern Women Data
{{< include southern-women.html >}}

## What do we get when we do CA?
:::: {.columns}
::: {.column width="50%"}
- Taken singly, the first dimension of CA suggests an **ordering**
  - Is this connected to centrality?
  - If so, what is the relation (if any) to existing two-mode centralities?*
- Take jointly, the first two dimensions of CA suggest a **grouping**
  - Connected to group cohesion?**
  - Affinity of actors for particular events?***
:::

::: aside
 *Bonacich 1991

 **Doreian 1979

 ***Faust 2005
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data Obtained from Simple CA of the Affiliation Matrix](Plots/ca-southern-women.png)
:::
::::

## What do we get when we do CA?

- To (try to) shed light on these questions I examine the cases of three relatively recent "accidental reinventions" of CA in the "network science" and "data science" (computer and information science) literatures:
  - The "Economic Complexity Index" (ECI)*
    - Relevant to the question of centrality
  - *SimRank* and Generalized Relational Similarity**
    - Relevant to the question of grouping
- The ECI case has been examined in detail in recent work***
- The second and third are new to this presentation
  - All shed light on what exactly we get when we use CA to analyze two mode data

::: aside
*Hidalgo & Hausmann 2009

**Jeh & Widom 2002; Kovacs (2010)

***van Dam et al. 2021; Mealy 2019
:::

## CA and Reflective Centrality

- Hidalgo & Hausmann (2009) introduced a "novel" way to rank nodes in two-mode networks
  - Their approach, called "reflective," was originally applied to country-by-product networks*

. . .

  - The method has since been shown to be applicable to other two-mode networks (like cultural networks)**
  - I introduce here using the more familiar sociological example of the **duality of persons and groups*****

::: aside
*Referred to as the Economic Complexity Index (ECI)

**Lizardo 2018

***Breiger 1974
:::

## CA and Reflective Centrality

- Notation:
  - A two-mode network composed of a set of persons $P$
  - Their affiliation relations to a set of groups $G$
  - Represented by an affiliation matrix $\mathbf{A}$

. . .

  - $\mathbf{A}$ is of dimensions $|P| \times |G|$ with people along the rows and groups across the columns
    - $|P|$ is the cardinality of the persons set  
    - $|G|$ is the cardinality of the group set

. . .

  - $\mathbf{A}$ has cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.

## CA and Reflective Centrality
- If we are going to compute the centrality of nodes in a two-mode network, the most natural place to start is with the good old **degree centrality***

- Given this, the degree-centrality of people is given by:

$$
C^R_p(1) = \sum_g a_{pg}
$$ {#eq-dp}

- And for groups:

$$
   C^R_g(1) = \sum_p a_{pg}
$$ {#eq-dg}

:::aside
*Faust 1997, eqs. 4 and 5
:::

## CA and Reflective Centrality

- Once we have the first-order quantities $C^R_p(1)$ and $C^R_g(1)$, we can compute "second-order centralities" $C^R(2)$ for both people and groups

- For people, these are given by:*

$$
C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
$$ {#eq-r2p}

- And for groups:**

$$   
C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
$$ {#eq-r2g}

:::aside
*People are central when the average sum of the size of the groups they belong to is large

**Groups are central when the average activity of the people who belong to them is large
:::

::: {.notes}
Equation @eq-r2p says "people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). Equation @eq-r2g says "groups are more central when the average activity of their members is large" (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number).
:::

## CA and Reflective Centrality

- Of course, once we have the second-order quantities $C^R_p(2)$ and $C^R_g(2)$, we can compute "third-order centralities" $C^R(3)$ for both people and groups

- For people, these are given by:

$$
   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
$$ {#eq-r3p}

- And for groups:

$$   
   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
$$ {#eq-r3g}

:::aside
*People are central when the average sum of the average activity of the members of the groups they belong to is large

**Groups are central when the average sum of the average size of the groups their members belong to is large
:::

::: {.notes}
Equation @eq-r3p says something like "people are more central when the average sum of the average activity of the members of the groups they belong to is large" (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number). Equation @eq-r3g says, "groups are more central when the average sum of the average size of the groups their members belong to is large." (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).
:::

## CA and Reflective Centrality
- More generally, Hidalgo & Hausmann show that we can define a **series** of reflective quantities* $C^R_p(4), C^R_p(5)...C^R_p(q)$ and $C^R_g(4), C^R_g(5)...C^R_g(q)$

- For people, these are given by:

$$   
C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
$$ {#eq-rip}

- And for groups:

$$   
C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
$$ {#eq-rig}

::: aside
*Whose verbal and substantive interpretation becomes more elusive as the number of iterations increases
:::

## CA and Reflective Centrality

- For people:
  - Even-numbered reflections measure centrality based on the **average size** of the groups a person belongs to* 
  - Odd-numbered reflections measure centrality based the **average number of memberships** of each member of the groups they belong to*

. . .

- For groups:
  - Even-numbered reflections measure centrality based on the **average number of memberships** of the members of the group*
  - Odd-numbered reflections measure centrality based on the **average size** of the group memberships of each member*

::: aside
*and the average of the average of the average of the average...
:::

## Reflection Trajectories for SWD
:::: {.columns}
::: {.column width="50%"}
![Person Reflection Trajectories](Plots/p-reflections.png)
:::

::: {.column width="50%"}
![Group Reflection Trajectories](Plots/g-reflections.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Person Reflection Trajectories](Plots/p-reflections.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Person Reflection Trajectories](Plots/g-reflections.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Scatter of standardized person reflection centralities and person scores on first CA dimension (r = 1.0)](Plots/p-ca-ref-corr.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Final Reflection Ordering versus CA
:::: {.columns}
::: {.column width="50%"}
![Scatter of standardized group reflection centralities and group scores on first CA dimension (r = 1.0)](Plots/g-ca-ref-corr.png)
:::

::: {.column width="50%"}
![Correspondence Plot of Southern Women Data](Plots/ca-southern-women.png)
:::
::::

## Method of Reflections versus CA

- So the method of reflections is an accidental re-invention of CA!*

- But *why* is it a reinvention of CA?
  - Short answer: 
    - The final reflection scores are the (second largest) eigenvalues of a matrix, just like CA!

- What's the (slightly) longer answer? 
  - A couple of things first:
    - $Dp$ is a $|P| \times |P|$ square degree matrix containing each person's $C_p^R(1)$ (activity) along the diagonals and zeroes in every other cell
    - $Dg$ is a $|G| \times |G|$ square degree matrix containing each group's $C_g^R(1)$ (size) along the diagonals and zeroes in every other cell

:::aside
*On the first dimension: Mealy 2019; van Dam et al. 2021
:::

## Method of Reflections versus CA
- So the method of reflections is an accidental re-invention of CA!*

- But *why* is it a reinvention of CA?

  - Short answer: 
    - The final reflection scores are the (second largest) eigenvalues of a matrix, just like CA!

- What's the (slightly) longer answer?

  - In the limit ($q \rightarrow \infty$), the iterative Hidalgo-Hausmann reflective centralities are the solution of the eigensystem:*

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
$$ {#eq-dam1}

$$
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
$$ {#eq-dam2}

::: aside
*van Dam et al. 2021
:::

## CA versus Bonacich
- Do the matrices whose eigenvalues yield the equilibrium reflective centralities look familiar to you?

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
$$

$$
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
$$

- They should, because the solutions to the following eigensystem give the Bonacich $(C^B)$ dual centralities for two-mode networks:*

$$
\left(AA^T\right)C^B_p = \lambda C^B_p
$$ {#eq-bon1}

$$
\left(A^TA\right)C^B_g = \lambda C^B_g
$$ {#eq-bon2}

:::aside
*Bonacich 1991
:::

## CA versus Bonacich
- So both Reflective and Bonacich centralities are eigenvectors of the person and group **projections** of the two-mode network

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\\
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\\
\left(AA^T\right)C^B_p = \lambda C^B_p
\\
\left(A^TA\right)C^B_g = \lambda C^B_g
$$

- But for reflective centralities the affiliation matrix and its transpose are pre-multiplied by the **inverse of the degree matrix of each mode***

:::aside
* van Dam et al. 2021
:::

## CA versus Bonacich
- So both Reflective and Bonacich centralities are eigenvectors of the person and group **projections** of the two-mode network

$$
\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
\\
\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
\\
\left(AA^T\right)C^B_p = \lambda C^B_p
\\
\left(A^TA\right)C^B_g = \lambda C^B_g
$$

- Similar to how each entry in the original affiliation matrix is weighted in CA performing the SVD!*

$$
Dp^{-1/2}ADg^{-1/2} = X_p \lambda X_g^T
$$

::: aside
**Faust 2005, eq. 7.8 
:::

## CA versus Bonacich
- We can gain some insight into the equivalence between CA and Reflections by expressing the degree-weighted projections in terms of each cell entry:*

$$
a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
$$ {#eq-mealy1}

  - In @eq-mealy1, the fraction's numerator is equal to one only when person $p$ and $p'$ are both members of $g$.

    - Summed across groups, this is the number of **shared groups** between $p$ and $p'$: 
      - The cell entries in the Bonacich matrix (raw projections)

    - But in @eq-mealy1 the numerator is *weighted* by the product of group size and each person's number of memberships (in contrast to Bonacich)

::: aside
*Mealy 2019, eq. 4
:::

## CA versus Bonacich
- We can gain some insight into the equivalence between CA and Reflections by expressing the degree-weighted projections in terms of each cell entry:*

$$
    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
$$ {#eq-mealy2}

  - In @eq-mealy2, the fraction's numerator is equal to one only when group $g$ and $g'$ both have $p$ as a member

    - Summed across persons, the number of **shared members** between $g$ and $g'$:
      - The cell entries in the Bonacich matrix (raw projections)

    - But in @eq-mealy2 the numerator is *weighted* by the product of group size and each person's number of memberships (in contrast to Bonacich)

::: aside
*Mealy 2019, eq. 4
:::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
![Bonacich Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- The CA re-ordered affiliation matrix is **block-diagonal** 
- Reveals (a well-known and storied*) grouping of persons and groups into two clusters 
![](Plots/doreian-sw.png)
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central events contain the most central people and central people attend the most central events:
  - $E1 = \{Laura, Brenda, Evelyn\}$
- In CA ordering central events are *not* the ones with most members! 
  - $\{E7, E8, E9\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central events contain the most central people and central people attend the most central events:
  - $E1 = \{Laura, Brenda, Evelyn\}$
- But they are the most central in the Bonacich ordering!
  - $\{E7, E8, E9\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Least central events contain many of the least central people:
  - $E11 = \{Helen, Flora, Olivia, \\ Nora\}$
- In CA ordering less central events are *not* the ones with the least members! 
  - $\{E1, E13, E14\}$
  - $E1$ is the most central!
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central people attend the most central events and central events are attended by central people:
  - $Frances = \{E3, E5, E4, E6, E8\}$
- Not necessarily those with the most memberships
  - $\{Theresa, Evelyn, Brenda\}$

:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- The least central person $\{Nora\}$ is not the one with the least memberships
  - Instead $\{Nora\}$ shares many memberships with less central people (those who attend the less central events)
  - $\{Flora,  Olivia, Sylvia, Kat.\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::

::: {.column width="50%"}
- First CA dimension is also an **ordering** so it counts as a centrality (a rank over the nodes)
- It is an ordering that uses a logic of grouping in order to rank
  - Tasks usually segregated in sna
- Central people attend the most central events and central events are attended by central people:
  - $E1 = \{Laura, Brenda, Evelyn\}$
- In CA ordering central events are *not* the ones with the most members! 
  - e.g., $\{E7, E8, E9\}$
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![Bonacich Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::
::: {.column width="50%"}
- The Bonacich re-ordered affiliation matrix is **upper triangular**
- Arranges persons and groups from top to bottom and left to right by activity and size respectively
  - More central people and groups at the upper right
  - Less central people and groups at the lower left
:::
::::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![Bonacich Re-ordered Affiliation Matrix](Plots/bon-reord.png)
:::
::: {.column width="50%"}
- Bonacich centrality is dual but it is "simple" dual
  - Central people attend well-attended events
  - Central events are attended by people who attend many events
- Centrality in one mode is a weighted *sum* of the other mode's centrality*

$$
    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
$$ {#eq-fau1}

$$
    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
$$ {#eq-fau2}
:::
::::

:::aside
*Faust 1997, p. 170
:::

## CA versus Bonacich
:::: {.columns}
::: {.column width="50%"}
![CA Re-ordered Affiliation Matrix](Plots/ca-reord.png)
:::
::: {.column width="50%"}
- CA centrality is dual but it is "generalized similarity" dual*
  - Central people attend central (not necessarily the most well-attended) events
  - Central events are attended by central (not necessarily the most active) people
- Some people calculate Bonacich centralities when they have this more "generalized similarity" idea of centrality in mind!
  - They should use CA instead
:::
::::
:::aside
*Kovacs 2010; Lizardo 2024
:::

## CA and Similarity
- Another way to gain insight into what CA is doing is by looking at the similarity of people and groups in the data

- As we all know,* the usual one mode projections are obtained via:

$$
S_p = AA^T
$$ {#eq-brei1}

$$
S_g = A^TA
$$ {#eq-brei1}

- Which can be thought of as **similarity matrices** based on the shared neighbors in the two mode network

::: aside
Breiger 1974
:::

## CA and Similarity
- Another way to gain insight into what CA is doing is by looking at the similarity of people and groups in the data

- But we can also obtain *other-mode degree weighted* versions of these similarity matrices*

$$
S^*_p = ADg^{-1}A^T
$$ {#eq-dam5}

$$
S^*_g = A^TDp^{-1}A
$$ {#eq-dam6}

::: aside
van Dam et al. 2021
:::

## CA and Similarity
:::: {.columns}
::: {.column width="50%"}
![Degree-weighted similarity matrix (persons)](Plots/ca-sim-p.png)
:::
::: {.column width="50%"}
![First eigenvalue of person similarity matrix against rank](Plots/ca-eigvec-p.png)
:::
::::

## CA and Similarity
:::: {.columns}
::: {.column width="50%"}
![Degree-weighted similarity matrix (groups)](Plots/ca-sim-g.png)
:::
::: {.column width="50%"}
![First eigenvalue of group similarity matrix against rank](Plots/ca-eigvec-g.png)
:::
::::

## CA and Similarity
:::: {.columns}
::: {.column width="30%"}
![](Plots/ca-sim-p.png)
![](Plots/ca-sim-g.png)

:::
::: {.column width="55%"}
![CA Correspondence Plot](Plots/ca-corr-plot.png)
:::
::::

## Bonacich and Similarity
:::: {.columns}
::: {.column width="30%"}
![](Plots/bon-sim-p.png)
![](Plots/bon-sim-g.png)

:::
::: {.column width="55%"}
![Bonacich Correspondence Plot](Plots/bon-corr-plot.png)
:::
::::

## CA and Generalized Relational Similarity
- Bonacich orders people according to attendance at central (most-well attended) events and events according to attendance by central (most active) people
- CA orders people according to attendance at **similarly central events** and events according to attendance by **similarly central people** 
  - This is precisely the definition of **generalized relational similarity** (GSR)*

. . .

- In that case proximity of nodes in the first CA dimension may reveal a type of GSR in two-mode networks
  - Let us see how that works in the case of two candidate two-mode GSRs
    - *SimRank* *
    - Similarity Weighted Correlation Distance**

:::aside
*Jeh & Widom 2002

**Kovacs 2010; Lizardo 2024
:::

## CA and SimRank

- To explore the link between CA and GSR we can compare CA with one independently developed approach for obtaining GSRs in two-mode networks:* 
  - *SimRank*

- According to *SimRank*:
  - People are *similar* if they belong to *similar* groups
  - Groups are *similar* if they share *similar* members

- In *SimRank* the similarity between two people is given by:

$$
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
$$ {#eq-simrank1}

:::aside
*Jeh & Widom 2002
:::

## CA and SimRank

- To explore the link between CA and GSR we can compare CA with one independently developed approach for obtaining GSRs in two-mode networks:* 
  - *SimRank*

- According to *SimRank*:
  - People are *similar* if they belong to *similar* groups
  - Groups are *similar* if they share *similar* members

- In *SimRank* the similarity between two groups is given by:

$$
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
$$ {#eq-simrank2}

:::aside
*Jeh & Widom 2002
:::

## CA and SimRank

- *SimRank*  for people:

$$
\small
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
$$

- *SimRank*  for groups:

$$
\small
S(g, g') = \frac{\alpha}{C_g^R(1)C_{g'}^R(1)}
    \sum_{i = 1}^{C_g^R(1)} \sum_{j = 1}^{C_{g'}^R(1)} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
$$

With:*

- $S(p, p) = 1$ and $S(g, g) = 1$ and $0 > \alpha < 1$

:::aside
*Jeh & Widom 2002
:::

::: {.notes}
In two-mode networks, SimRank scores can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation @eq-simrank1 using baseline values. Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset. We then plug those values into equation @eq-simrank2, then loop back to equation @eq-simrank1 with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations
:::

## CA and SimRank
:::: {.columns}
::: {.column width="50%"}
![SimRank Re-ordered Affiliation Matrix](Plots/sr-plot-reord.png)
:::

::: {.column width="50%"}
![Simrank Similarity Matrix Eigenvalue Plot (Persons)](Plots/sr-plot-eigen-p.png)
:::
::::

## CA and SimRank
:::: {.columns}
::: {.column width="50%"}
![SimRank Re-ordered Affiliation Matrix](Plots/sr-plot-reord.png)
:::

::: {.column width="50%"}
![Simrank Similarity Matrix Eigenvalue Plot (Groups)](Plots/sr-plot-eigen-g.png)
:::
::::

## CA and SimRank
:::: {.columns}
::: {.column width="50%"}
![SimRank Re-ordered Affiliation Matrix](Plots/sr-plot-reord.png)
:::

::: {.column width="30%"}
![Correlation between CA and SimRank Similarity Matrix Eigenvectors](Plots/sr-corr-scatter.png)
:::
::::

## CA and Generalized Relational Similarity

- I recently proposed a measure of generalized relational similarity (GSR) metric for nodes in two-mode networks by "tweaking" a previous approach based on the modified correlation distance*

- For pairs of people the GSR is given by:

$$
\begin{equation}
\scriptsize
    S(p, p') = 
    \frac{
    (A_{p\bullet} - \bar{A}_{p\bullet})
    \mathbf{S(G)}
    (A_{p'\bullet} - \bar{A}_{p'\bullet})^T
    }
    {
    \sqrt{
    (A_{p\bullet} - \bar{A}_{p\bullet})
    \mathbf{S(G)}
    (A_{p\bullet} - \bar{A}_{p\bullet})^T
    }
    \sqrt{
    (A_{p'\bullet} - \bar{A}_{p'\bullet})
    \mathbf{S(G)}
    (A_{p'\bullet} - \bar{A}_{p'\bullet})^T
        }
    }
\end{equation}
$${#eq-kovacs1}

- And for groups:

$$
\begin{equation}
\scriptsize 
    S(g, g') = 
    \frac{
    (A_{g\bullet} - \bar{A}_{g\bullet})
    \mathbf{S(P)}
    (A_{g'\bullet} - \bar{A}_{g'\bullet})^T
    }
    {
    \sqrt{
    (A_{g\bullet} - \bar{A}_{g\bullet})
    \mathbf{S(P)}
    (A_{g\bullet} - \bar{A}_{g\bullet})^T
    }
    \sqrt{
    (A_{g'\bullet} - \bar{A}_{g'\bullet})
    \mathbf{S(P)}
    (A_{g'\bullet} - \bar{A}_{g'\bullet})^T
        }
    }
\end{equation}
$${#eq-kovacs2}

:::aside
*Lizardo 2024; Kovacs 2010
:::

## CA and Generalized Relational Similarity

- With the base similarities between people given by:

$$
\mathbf{S(P)}_{pp'} = \frac{AA_{pp'}}{\sqrt{AA_{pp}AA_{p'p'}}}
$${#eq-lizardo1}

- And the base similarities for groups:

$$
\mathbf{S(G)}_{gg'} = \frac{AA_{gg'}}{\sqrt{AA_{gg}AA_{g'g'}}}
$${#eq-lizardo2}

- Namely, the cosine distances between people and groups based on the corresponding cell entries of the one-mode projections*

:::aside
*Lizardo 2024
:::

## CA and Generalized Relational Similarity
:::: {.columns}
::: {.column width="50%"}
![GRS Re-ordered Affiliation Matrix](Plots/grs-plot-reord.png)
:::

::: {.column width="50%"}
![GRS Similarity Matrix Eigenvalue Plot (Persons)](Plots/grs-plot-eigen-p.png)
:::
::::

## CA and Generalized Relational Similarity
:::: {.columns}
::: {.column width="50%"}
![GRS Re-ordered Affiliation Matrix](Plots/grs-plot-reord.png)
:::

::: {.column width="50%"}
![GRS Similarity Matrix Eigenvalue Plot (Groups)](Plots/grs-plot-eigen-g.png)
:::
::::

## CA and Generalized Relational Similarity
:::: {.columns}
::: {.column width="50%"}
![GRS Re-ordered Affiliation Matrix](Plots/grs-plot-reord.png)
:::

::: {.column width="30%"}
![Correlation between CA and GRS Matrix Eigenvectors](Plots/grs-corr-scatter.png)
:::
::::

## Conclusions
- To paraphrase John Martin paraphrasing Levi-Strauss:
  - There is only *one analysis* in two-mode network analysis and it is CA
  - Rediscovered many times by researchers pursuing seemingly incompatible taks

. . .

- Contra skeptics we can do two-mode "sna" when we do CA
  - Rank nodes against a meaningful centrality metric
  - Detect equivalences (generalized similarities) between node clusters
- These are some of the standard "positional" tasks that take up much sna

## Conclusions
- To paraphrase John Martin paraphrasing Levi-Strauss:
  - There is only *one analysis* in two-mode network analysis and it is CA
  - Rediscovered many times by researchers pursuing seemingly incompatible taks

. . .

- We can also go do more than visualization and joint display
- The first CA dimension particularly represents a seldom noted **duality of ranking and equivalence**
  - Ranking via similarity within clusters
  - Bonacich by contrast is a "pure" centrality ranking

## Conclusions
- To paraphrase John Martin paraphrasing Levi-Strauss:
  - There is only *one analysis* in two-mode network analysis and it is CA
  - Rediscovered many times by researchers pursuing seemingly incompatible taks

. . .

- Using CA for this kind of "spectral clustering" and "spectral ranking" is not common but feasible
  - Another approach (similar to blockmodeling):
    - Use CA for spectral clustering
    - Then use **specific CA** (on sub-tables defined by the eigengap heuristic) to split clusters into further clusters

## Thanks!

[https://github.com/olizardo/correspondence-analysis-two-mode-networks](https://github.com/olizardo/correspondence-analysis-two-mode-networks)

olizardo@soc.ucla.edu