### 2024-07-12 analysis.qmd
Move around some things

---
title: "Two Mode Networks and Relational Similarity"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
---

# Setup
```{r setup}
   knitr::opts_chunk$set(include = TRUE, 
                       echo = FALSE, 
                       warning = FALSE, 
                       message = FALSE)
   library(cowplot)
   library(here)
   library(patchwork)
   library(dplyr)
   library(factoextra)
   library(FactoMineR)
   library(ggbump)
   library(ggcorrplot)
   library(ggplot2)
   library(ggpubr)
   library(ggrepel)
   library(igraph)
   library(kableExtra)
   library(pals)
   library(patchwork)
```

# Southern Women Data Table
```{r HTML data table}
   source(here("Functions", "sw.dat.R"))
   A <- A[sort(rownames(sw.dat())), ]
   kbl(A, align = "c", format = "html") %>% 
      kable_minimal(full_width = F, font_size = 26) %>% 
      row_spec(1:18, extra_css = "border-bottom: 0px;line-height: 20px;") %>% 
      save_kable(here("southern-women.html"))
```

# Analysis
```{r CA of affiliation matrix}
   ca.res <- CA(A, graph = FALSE)
   plot.dat <- data.frame(rbind(ca.res$row$coord, ca.res$col$coord)[, 1:2])
   plot.dat <- cbind(plot.dat, mode = c(rep(1, 18), rep(2, 14)), names = rownames(plot.dat))
   p <- ggplot(plot.dat, aes(x = Dim.2, y = Dim.1*-1, 
                             color = factor(mode), label = names))
   p <- p + geom_vline(xintercept = 0, color = "gray")
   p <- p + geom_hline(yintercept = 0, color = "gray")
   p <- p + geom_text_repel(size = 5, max.overlaps = 50)
   p <- p + theme_minimal() 
   p <- p + theme(legend.position = "none",
                  axis.text = element_text(size = 20),
                  axis.title = element_text(size = 24)
                  )
   p <- p + labs(x = "Second Dimension", y = "First Dimension")
   p <- p + scale_color_manual(values = c('red', 'blue'))
   png(file = here("Plots", "ca-southern-women.png"), width=600, height=625)
   p
   dev.off()
```

```{r Reflections Bump Chart (Persons)}
   source(here("Functions", "reflections.R"))
   source(here("Functions", "ref.long.dat.R"))
   evens <- function(x) subset(x, x %% 2 == 0)
   odds <- function(x) subset(x, x %% 2 != 0)
   a <- reflections(A, iter = 26)
   p.e.r <- a$p.r[, evens(1:20)]
   g.e.r <- a$g.r[, evens(1:20)]
   b <- ref.long.dat(a = p.e.r, b = g.e.r, max.iter = 26)
   plot.dat <- b$person
   p <- ggplot(data = filter(plot.dat, n.iter < 18),
               mapping = aes(x = n.iter, y = ref, color = person))
   # geoms
   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
   p <- p + geom_point(size = 5)
   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
                                    aes(x = n.iter, label = person), 
                                    size = 5, nudge_x = -3)
   p <-    p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 18), 
                              aes(x = n.iter, label = person), 
                              size = 5, nudge_x = 0.01)

   # fixings
   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
   p <- p + scale_x_continuous(limits = c(-2, 20), breaks = evens(2:16))
   p <- p + labs(x = "Reflection")
   p <- p + scale_y_reverse()
   p <- p + theme(legend.position = "none",
        panel.grid.major = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 16)
        )
   set.seed(145)
   c <- sample(cols25(25), 18, replace = FALSE)
   p1 <- p + scale_color_manual(values=as.vector(c))
   png(file = here("Plots", "p-reflections.png"), width=600, height=625)
   p1
   dev.off()
```

```{r Reflections Bump Chart (Groups)}
   plot.dat <- b$group
   p <- ggplot(data = filter(plot.dat, n.iter < 14),
               mapping = aes(x = n.iter, y = ref, color = group))
   #geoms
   p <- p + geom_bump(linewidth = 1.1, smooth = 8)
   p <- p + geom_point(size = 5)
   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 2), 
                                    aes(x = n.iter, label = group), 
                                    size = 6, nudge_x = -1)
   p <- p + geom_text_repel(data = filter(plot.dat, n.iter == 14), 
                              aes(x = n.iter, label = group), 
                              size = 6, nudge_x = -0.01)
   #fixings
   p <- p + theme_minimal_grid(font_size = 14, line_size = 0)
   p <- p + scale_x_continuous(limits = c(1, 14), breaks = evens(2:12))
   p <- p + labs(x = "Reflection")
   p <- p + scale_y_reverse()
   p <- p + theme(legend.position = "none",
        panel.grid.major = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 16)
        )
   set.seed(145)
   c <- sample(cols25(25), 18, replace = FALSE)
   p2 <- p + scale_color_manual(values=as.vector(c))
   png(file = here("Plots", "g-reflections.png"), width=600, height=625)
   p2
   dev.off()
```

```{r CA versus Reflections Scatter}
   tm <- TwoModeCA(A)
   mr <- reflections(A, iter = 26)
   p.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.r[, 1], 
                                  r = mr$p.s[, 22]))
   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
   p <- p + geom_point(size = 3, color = "blue") 
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "18th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()

   p.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1], 
                                  r = mr$g.s[, 16]))
   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
   p <- p + geom_point(size = 3, color = "red") 
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), 
                            color = "blue", size = 6)
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "14th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()
```

```{r CA via SVD of two mode network}
   Rp <- diag(sqrt(rowSums(A)))
   Rc <- diag(sqrt(colSums(A)))

   svd.res <- svd(Rp%*%A%*%Rc)

   t <- sum(rowSums(A))
   r.rp <- t/rowSums(A)
   r.cp <- t/colSums(A)
   s.rp.cp <- sqrt(c(r.rp, r.cp))
   tv <- t(svd.res$v)
   x <- c(svd.res$u[, 3]*r.rp, tv[3, ]*r.cp) 
   y <- c(svd.res$u[, 2]*r.rp, tv[2, ]*r.cp)

   d <- data.frame(x = x, y = y, l = c(rownames(A), colnames(A)))
   ggscatter(d, x = "x", y = "y", label = "l", repel = TRUE)
```

```{r CA versus Bonacich Re-ordered Affiliation Matrices}
   a <- TwoModeCA(A)
   b <- a$ca.A
   c <- a$bon.A
   png(file = here("Plots", "ca-reord.png"), width=500, height=500)
   b
   dev.off() 
   png(file = here("Plots", "bon-reord.png"), width=500, height=500)
   c
   dev.off() 
   png(file = here("Plots", "ca-v-bon-reord.png"), width=1000, height=700)
      (b | c) + plot_annotation(tag_levels = "a") &
      theme(plot.tag = element_text(size = 20))
   dev.off() 
```

```{r CA Eigenvector Plot}
    tmCA.res <- TwoModeCA(A, k.r = 2, k.c = 2)
    a <- tmCA.res$eigvec.plot.r + theme(axis.text = element_blank())
    b <- tmCA.res$eigvec.plot.c 
    c <- tmCA.res$eigval.plot.r
    d <- tmCA.res$eigval.plot.c
    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)

   png(file = here("Plots", "ca-eigvec-p.png"), width=500, height=500)
   a
   dev.off() 
   png(file = here("Plots", "ca-eigvec-g.png"), width=500, height=500)
   b
   dev.off() 
    png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
        (a | b) + plot_annotation(tag_levels = "a") &
      theme(plot.tag = element_text(size = 20))
    dev.off() 
```

```{r Similarity and Correspondence Plot (CA)}
    tmCA.res <- TwoModeCA(A, d.r = 2, d.c = 2, k.r = 5, k.c = 5)
    a <- tmCA.res$ca.sim.plot.r + theme(axis.text.x = element_blank())
    b <- tmCA.res$ca.sim.plot.c + theme(axis.text.x = element_blank())
    c <- tmCA.res$ca.corr.plot

    png(file = here("Plots", "ca-sim-p.png"), width=600, height=625)
    a + theme(axis.text.y = element_text(size = 18))
    dev.off()

    png(file = here("Plots", "ca-sim-g.png"), width=600, height=625)
    b + theme(axis.text.y = element_text(size = 22))
    dev.off()

    png(file = here("Plots", "ca-corr-plot.png"), width=600, height=625)
    c
    dev.off()
```

```{r Similarity and Correspondence Plot (Bonacich)}
    tmCA.res <- TwoModeCA(A, d.r = 2, d.c = 2, b.r = 4, b.c = 4)
    a <- tmCA.res$bon.sim.plot.r + theme(axis.text.x = element_blank())
    b <- tmCA.res$bon.sim.plot.c + theme(axis.text.x = element_blank())
    c <- tmCA.res$bon.corr.plot
    png(file = here("Plots", "bon-sim-p.png"), width=600, height=625)
      a + theme(axis.text.y = element_text(size = 16))
    dev.off()

    png(file = here("Plots", "bon-sim-g.png"), width=600, height=625)
      b + theme(axis.text.y = element_text(size = 16))
    dev.off()

    png(file = here("Plots", "bon-corr-plot.png"), width=600, height=625)
      c
    dev.off() 
```

```{r SimRank Plot}
   source(here("Functions", "TwoModeSR.R"))
   tmSR.res <- TwoModeSR(A)
    a <- tmSR.res$eigvec.plot.r 
    b <- tmSR.res$eigvec.plot.c 
    c <- tmSR.res$eigval.plot.r
    d <- tmSR.res$eigval.plot.c
    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    e <- tmSR.res$A.plot

   dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1]),
                       sr = tmSR.res$ev.r1,
                       lab = rownames(A))
   dat.c <- data.frame(ca = as.numeric(tm$ca.eigvec.c[,1])*-1,
                       sr = tmSR.res$ev.c1,
                       lab = colnames(A))
   cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
   g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
                  label = "lab", repel = TRUE, 
                  add = "reg.line", 
                  add.params = list(color = "gray80", 
                                    linewidth = 0.5, linetype = 2)) +
                  annotate("text", x = -.2, y = .3, 
                           label = paste("r = ", cor), size = 6) 
   cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
   h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
                  label = "lab", repel = TRUE, 
                  add = "reg.line",
                  add.params = list(color = "gray80", 
                                    linewidth = 0.5, linetype = 2)) +
                  annotate("text", x = -.2, y = .3, 
                           label = paste("r = ", cor), size = 6)

   png(file = here("Plots", "sr-plot-eigen-p.png"), width=600, height=625)
   a
   dev.off()
   png(file = here("Plots", "sr-plot-eigen-g.png"), width=600, height=625)
   b
   dev.off()
   png(file = here("Plots", "sr-plot-reord.png"), width=600, height=625)
   e
   dev.off()
   png(file = here("Plots", "sr-corr-scatter.png"), width=350, height=625)
   g / h
   dev.off()
```

```{r GRS Plot}
   source(here("Functions", "TwoModeGRS.R"))
   tmSR.res <- TwoModeGRS(A)
    a <- tmSR.res$eigvec.plot.r 
    b <- tmSR.res$eigvec.plot.c 
    c <- tmSR.res$eigval.plot.r
    d <- tmSR.res$eigval.plot.c
    a <- a + inset_element(c, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    b <- b + inset_element(d, left = 0.01, right = 0.4, bottom = 0.55, top = 1)
    e <- tmSR.res$A.plot

   dat.r <- data.frame(ca = as.numeric(tm$ca.eigvec.r[,1])*-1,
                       sr = tmSR.res$ev.r1,
                       lab = rownames(A))
   dat.c <- data.frame(ca = as.numeric(tm$ca.eigvec.c[,1])*-1,
                       sr = tmSR.res$ev.c1,
                       lab = colnames(A))
   cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)
   g <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
                  label = "lab", repel = TRUE, 
                  add = "reg.line", 
                  add.params = list(color = "gray80", 
                                    linewidth = 0.5, linetype = 2)) +
                  annotate("text", x = -.2, y = .3, 
                           label = paste("r = ", cor), size = 6) 
   cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)
   h <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
                  label = "lab", repel = TRUE, 
                  add = "reg.line",
                  add.params = list(color = "gray80", 
                                    linewidth = 0.5, linetype = 2)) +
                  annotate("text", x = -.2, y = .3, 
                           label = paste("r = ", cor), size = 6)

   png(file = here("Plots", "grs-plot-eigen-p.png"), width=600, height=625)
   a
   dev.off()
   png(file = here("Plots", "grs-plot-eigen-g.png"), width=600, height=625)
   b
   dev.off()
   png(file = here("Plots", "grs-plot-reord.png"), width=600, height=625)
   e
   dev.off()
   png(file = here("Plots", "grs-corr-scatter.png"), width=350, height=625)
   g / h
   dev.off()
```