### 2025-01-29 highlights.tex

\documentclass{article}
\begin{document}

\section*{Highglights}

\begin{itemize}
  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
  \item CA can extract latent positional information on people and groups, akin to recent work on ``generalized relational similarity" in two-mode networks. 
  \item The Bonacich dual centrality is useful for highlighting core-periphery patterns in the two-mode case in contrast to CA scores which are more appropriate to revealing the latent group structure 
\end{itemize}

\end{document}