### 2025-02-01 main-R1.tex

%% 
%% Copyright 2019-2020 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-dc documentclass for 
%% double column output.

%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
\documentclass[a4paper,fleqn]{cas-sc}

%\usepackage[numbers]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[authoryear,longnamesfirst]{natbib}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{longtable}

%%%Author definitions
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{WGM}
\tsc{QE}
\tsc{EP}
\tsc{PMS}
\tsc{BEC}
\tsc{DE}
%%%

% Uncomment and use as if needed
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}[theorem]{Lemma}
%\newdefinition{rmk}{Remark}
%\newproof{pf}{Proof}
%\newproof{pot}{Proof of Theorem \ref{thm}}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}

% Short title
\shorttitle{Correspondence Analysis of Two-Mode Networks}

% Short author
%\shortauthors{O. Lizardo}

% Main title of the paper
\title [mode = title]{The Correspondence Analysis of Two-Mode Networks Revisited}                      
% Title footnote mark
% eg: \tnotemark[1]
%\tnotemark[1,2]

% Title footnote 1.
% eg: \tnotetext[1]{Title footnote text}
% \tnotetext[<tnote number>]{<tnote text>} 
%\tnotetext[1]{}

%\tnotetext[2]{.}

% First author
%
% Options: Use if required
% eg: \author[1,3]{Author Name}[type=editor,
%       style=chinese,
%       auid=000,
%       bioid=1,
%       prefix=Sir,
       %orcid=0000-0002-5405-3007,
%       facebook=<facebook id>,
%       twitter=<twitter id>,
%       linkedin=<linkedin id>,
%       gplus=<gplus id>]
%\author[1]{Omar Lizardo}[orcid=0000-0002-5405-3007]

% Corresponding author indication
%\cormark[1]

% Footnote of the first author
%\fnmark[1]

% Email id of the first author
%\ead{olizardo@soc.ucla.edu}

% URL of the first author
%\ead[url]{http://olizardo.bol.ucla.edu/}

%  Credit authorship
%\credit{Conceptualization of this study, Methodology, Software}

%Address/affiliation
%\affiliation[1]{organization={University of California, Los Angeles},
   % addressline={264 Haines Hall}, 
   % city={Los Angeles},
    %citysep={},  Uncomment if no comma needed between city and postcode
    % postcode={90095}, 
    % state={CA},
    % country={USA}}

 %Corresponding author text
%\cortext[cor1]{Corresponding author}
%\cortext[cor2]{Principal corresponding author}

%Footnote text
%\fntext[fn1]{This paper was written while the author was partially supported by National Science Foundation grant HNDS-R \#2214215.}
 % author as well.}
%\fntext[fn2]{Another author footnote, this is a very long footnote and
 % it should be a really long footnote. But this footnote is not yet
 % sufficiently long enough to make two lines of footnote text.}

% For a title note without a number/mark
%\nonumnote{This note has no numbers. In this work we demonstrate $a_b$
%  the formation Y\_1 of a new type of polariton on the interface
 % between a cuprous oxide slab and a polystyrene micro-sphere placed
 % on the slab.
 % }
% Here goes the abstract
\begin{abstract}
This paper reconsiders the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting aspects that have not been previously emphasized, going beyond the use of CA as a visualization tool. It argues that CA can be used to compute a dual centrality score on both modes, related to but not mathematically equivalent to the \citet{bonacich1991simultaneous} two-mode centrality. This ``reflective'' centrality score connects the use of CA in two-mode network analysis to its use in other disciplines as a method of ordination. Additionally, I show CA can extract latent positional information on people and groups, with similarities to recent work on ``generalized relational similarity" in two-mode networks. Thus, CA can be used for indirect community or subgroup identification in two-mode networks, connecting to its use in some disciplines as a clustering method. The paper demonstrates these applications of CA, comparing them with the Bonacich dual centrality scores, and provides guidance on using CA for structural analysis in network analysis.
\end{abstract}

% Use if graphical abstract is present
% \begin{graphicalabstract}
% \includegraphics{figs/grabs.pdf}
% \end{graphicalabstract}

%Research highlights
\begin{highlights}
  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
\end{highlights}

% Keywords
% Each keyword is separated by \sep
\begin{keywords}
Correspondence Analysis \sep Generalized Relational Similarity \sep Duality \sep Two-mode networks \sep Centrality 
\end{keywords}

\maketitle
\newpage
\section{Introduction} \label{sec:intro}
While not particularly common, using Correspondence Analysis---hereafter CA---in analyzing two-mode network data has had a somewhat rocky career in the social networks literature. Initially criticized by \citet{borgatti1997network} as a relatively limited and perhaps even inapplicable tool, CA has found various proponents who see it as an important component of the SNA arsenal for two-mode network data analysis, particularly regarding its ability to economically provide synoptic (e.g., ``joint'') graphical representations of structural connectivity patterns across the two-modes \citep{roberts2000correspondence, breiger2000tool, faust2005using}, with the primary aim being to use CA---or its variants like Multiple Correspondence Analysis (MCA)---to ``visually explore'' such networks \citep{d2014use}. This paper contributes to the stream of previous work applying CA to analyze two-mode data. Nevertheless, it departs from the already-mentioned previous efforts in highlighting aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before, moving beyond the focus on data visualization.

Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}. With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general centrality score connecting the use of CA in two-mode network analysis to way CA features in some disciplines also concerned with two-mode data (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities. In social network analysis, this applies to conceptualizing the venerable duality between people and the groups they join \citep{breiger1974duality}.

In addition to its capacity to reveal latent ordinal structure in two-mode data, I also argue that CA can also be used to extract latent \textit{positional} information on people and groups. Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups). Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks connecting to how CA is used in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}, namely, the discovery of sociometrically similar subsets of entities in affiliation networks. In this way, CA emerges as a tool central to the usual network-analytic tasks and not just as a visualization or data-summarization tool.

\subsection{Organization of the Paper} \label{subsec:org}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA to produce a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes. This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead \citep{everett2013dual}. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

\section{Previous uses of CA and MCA in Two-Mode Network Data Analysis}

In this section, I consider previous contributions regarding the usefulness and limitations of CA as a methodological tool for the analysis of two-mode network data. For reasons of space and relevance, therefore my discussion is limited in two ways. First, I focus attention on (1) pieces discussing primarily methodological aspects of CA and (2) with regard to the analysis of two-mode network data or affiliation networks.\footnote{Naturally, I also exclude the rich literature that has used methods other than CA to analyze affiliation networks.} There are, of course, various substantive applications of CA to the analysis of social networks, as well as substantive and methodological applications of CA to the analysis of one-mode network data; these will not be part of the discussion here.

With these restrictions in mind, the literature on CA and two-mode networks can be divided into two streams. The first contributions focus on the usefulness of CA as a tool for the visualization or ``joint display'' of the two sets of entities in an affiliation network. Secondly, some contributions focus on the similarities and differences between CA and factoring---those based on a spectral decomposition of a suitable matrix---techniques for assigning scores to the nodes of a two-mode network, where the scores exploit the duality between the two sets of entities and could be, under some circumstances, interpreted as ``centrality'' scores.

\subsection{Previous Work on CA and Two-Mode Networks}

One of the earliest considerations of CA as a data-analytic tool for two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks. In that paper, \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people by the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique. According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (square quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation (in the least-squared error) sense of the similarities between row and column objects.  Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis. Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of the affiliation matrix---e.g., the Singular Value Decomposition (SVD). As we will see later, both can also be thought of as based on the spectral decomposition of the dual projection matrices.

As we will see later (section~\ref{subsec:refeigen}), the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated. Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight and describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques and for events, their likelihood of being attended by the most representative members of them 
\citep[164]{bonacich1991simultaneous}.

In an influential paper on centrality in two-mode networks, \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA. Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data. In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the event she attends; for events, their scores are given by the sum of the eigenvector centralities of its members. Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship was obtained between the CA scores corresponding to persons and groups. The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event. \citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell $a_{pg}$ is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event ($\sqrt{\sum_p a_{pg}\sum_g a_{pg}}$), something that she further clarify in a subsequent contribution \citep{faust2005using}.

\citet{borgatti1997network} provided another early and influential in-depth consideration of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect. The discussion was placed in a section of the paper titled ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as a visualization technique. According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events.'' Conversely, points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as a visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean, they would mislead interpreters who would try to impose such a metric to make sense of them.

In a paper published shortly thereafter \citet{roberts2000correspondence}, responded to some of the criticisms that \citet{borgatti1997network} levied at the use of CA for analyzing two-mode networks in their original piece. First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data and that, therefore, the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix. \citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and the event points in the two-dimensional representation and one that used all thirteen dimensions. Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events). Thus, the point distances are indeed interpretable as a kind of degree-weighted similarity between objects in each mode.

Perhaps the most thorough discussion of the uses of CA as a tool for the ``joint display'' of affiliation networks is that provided by \citet{faust2005using}. According to \citet[118]{faust2005using}, previous debates regarding the usefulness---or lack thereof---of CA as a visualization tool were marred by imprecision and lack of ``formal specification'' regarding the exact substantive meaning of the scores obtained via the procedure and a related lack of sustained discussion as to the exact ``relationship between the model and the input data.'' Faust's basic point is that CA provides a way to both place actors and events in an affiliation network in a joint space such that the inter-point distances are interpretable and such that the link to the original data (e.g., the affiliation network matrix) is transparent.

\citet{faust2005using} makes several contributions in this piece. First, Faust specifies the exact matrix that is subject to the singular value decomposition when CA is applied directly to the affiliation matrix of a two-mode network $\mathbf{A}$; namely, a scaled version where $\mathbf{A}$ is pre-multiplied by a diagonal matrix containing the square root of its row sums and post-multiplied by a diagonal matrix containing the square root of its column sums (equivalent to dividing every non-zero entry of $a_{pg}$ by the square root of the product of the number of memberships of person $p$ and the number of members of group $g$ before applying the SVD. Second, Faust specifies the exact relationship linking the resulting row and column CA scores and the original affiliation matrix input, providing the ``reconstruction equation" that takes us from the left and right singular vectors and eigenvalues to the former. Thirdly, Faust clarifies the duality linking the row and column scores, noting that the scores for the people are (activity) weighted sums of the scores that the groups they join, and the scores for the groups are (group size) weighted sums of the scores of their members.

Finally, \citet[128ff]{faust2005using} clarifies the link between the CA scores---and different normalizations thereof---assigned to persons and events in an affiliation network on each dimension and (same node-set) inter-point distance in a joint space. Like \citet{roberts2000correspondence} before her, Faust specifies that in the low (usually two) dimensional representation, inter-point distances are best (least squares) approximations of the chi-square distance between the membership profiles of pairs of people (or groups), where the profile of a given person-row or group-column is equal to the non-zero entries of the affiliation matrix divided by the corresponding row or column sum (people's activities and group sizes).\footnote{Note that because the chi-square distances are defined on these degree-weighted entries, they define a kind of degree-weighted similarity between pairs of people and pairs of groups, but not an exact similarity score as would be obtained, for instance, by using a criterion such as proportion matching \citep[208]{everett2013dual}.}

Worth mentioning in this regard is an intriguing recent contribution by \citet{d2014use}, who convincingly advocate for using Multiple Correspondence Analysis (MCA) for the joint display of persons and groups in affiliation networks. The key difference between CA and MCA is that MCA works with a transformed version of the affiliation matrix, in which the rows are treated as ``cases'' and the columns as categorical attributes of the case, where each possible value represents a modality that the column object can take with respect to the cases. In the case of binary network data, the affiliation matrix is transformed into a new matrix $\mathbf{Z}$ with the same number of rows (persons) as the original and with twice the number of columns representing either membership ($z_{pg+} = 1$) or non-membership ($z_{pg-} = 1$) in that a group. The MCA scores are obtained via SVD of a centered and doubly normalized version of $\mathbf{Z}$. \citet{d2014use} discuss various advantages of the MCA graphical representation of two-mode networks, including the capacity to produce interpretable joint displays of actors and events and the ability to incorporate exogenous actor and event attributes as supplementary variables. However, because what follows bears more directly on the use of simple CA for two-mode network analysis, this discussion is less relevant.

\section{Reflective Node Scores in Two-Mode Networks} \label{sec:ref2mode}
In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of assigning meaningful scores to nodes in a two-mode network using what they called at the time a ``reflective'' approach. Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products).\footnote{For a state-of-the-art review of the various substantive applications of the economic complexity approach see \citet{hidalgo2021economic}.} Subsequent work shows that there is no logical connection between the formal method and this particular application since the approach proposed can be used to analyze any two-mode data matrix. As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}. This is consistent with the spirit of the original approach, recently described by \citet[92]{hidalgo2021economic} as centered on the ``\textit{duality} between economic inputs and outputs'' (italics mine).

If we are going to assign substantively meaningful scores to nodes in a two-mode network, the most natural place to start is with the good old \textit{degree centrality} \citep{faust1997centrality}. A two-mode network composed of a set of people $P$ and their affiliation relations to a set of groups $G$ can be represented by an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ with people along the rows and groups across the columns, where $|P|$ is the cardinality of the people set and $|G|$ is the cardinality of the group set, with cell entries $a_{pg}= 1$ if person \textit{p} is affiliated with group \textit{g} and $a_{pg}= 0$ otherwise.

Given this, the degree-centrality of people is given by:

\begin{equation}
    C^R_p(1) = \sum_g a_{pg}
    \label{eq:R1_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(1) = \sum_p a_{pg}
  \label{eq:R1_g}
\end{equation}

That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups.

As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order'' quantities $C^R(2)$ for both people and groups using the \textit{averaged} first-order centralities of the entities in the other mode they are connected to.

For people, these are given by:

\begin{equation}
    C^R_p(2) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(1)
    \label{eq:R2_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(2) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(1)
    \label{eq:R2_g}
\end{equation}

While Equation~\ref{eq:R1_p} assigns a high value to people who belong to a lot of groups, Equation~\ref{eq:R2_p} assigns a high value to people who, on average, belong to large groups (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). In the same way, while Equation~\ref{eq:R1_g} assigns a high value to groups that have lots of members, Equation~\ref{eq:R1_g} assigns a high value to groups that, on average, have members who themselves have lots of memberships (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number). Of course, we can keep on going and define ``third-order'' reflections.

For the people, these are given by:

\begin{equation}
   C^R_p(3) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(2)
   \label{eq:R3_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(3) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(2)
   \label{eq:R3_g}
\end{equation}

As we saw, while for people, Equation~\ref{eq:R2_p} measured the averaged size of the groups they join, Equation~\ref{eq:R3_p} assigns a high value to people who, on average, belong to groups who are themselves attended by highly active members (e.g., whenever $a_{pg} = 1$ and $C^R_g(2)$ is a big number). In the same way, while Equation~\ref{eq:R2_g} assigns a high value to groups whose members have lots of memberships, Equation~\ref{eq:R3_g} assigns a high value to groups that, on average, have members who themselves (also on average) belong to large groups (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).

Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes). On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to). In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity). On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size). While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).

More generally, \citet{hidalgo2009building} show that we can define a series of reflective quantities for people and groups (whose verbal and substantive interpretation becomes more complex as the number of iterations increases).

For people, these are given by:

\begin{equation}   
    C^R_p(q) = \frac{1}{C^R_p(1)}\sum_g a_{pg}C^R_g(q-1) 
   \label{eq:Rq_p}
\end{equation}

And for groups:

\begin{equation}
   C^R_g(q) = \frac{1}{C^R_g(1)}\sum_p a_{pg}C^R_p(q-1)
   \label{eq:Rq_g}
\end{equation}

Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$. Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
    \caption{Southern Women data matrix with rows and columns ordered as in the original DGG table.}
    \label{fig:sw}
\end{figure}

\subsection{Empirical Example} \label{subsec:ex}
\subsubsection{Southern Women Data}
The following analyses use the classic ``Southern Women'' data \citep{davis1941} (DGG) as the running example. This data is doubly appropriate for showcasing the relevance of methods that exploit the duality between two sets of entities, as it was used in Breiger's \citeyearpar{breiger1974duality} classic paper. Moreover, Southern Women is a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts. The list has only gotten much longer in the last two decades since Freeman's methodological ``meta-analysis'' \citep[e.g.,][among many others]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, kovacs2010generalized, brusco2011analysis, borgatti2014analyzing, everett2013dual, lerner2022dynamic, batagelj2022analysis, lizardo2024two}. Despite this long list of studies of the Southern Women data, we will see that novel insights can still be garnered from this old data source.

Despite its undisputed pedigree as a methodological touchstone in studies of duality and dual constitution in two-mode networks, a key source of confusion in the mini-literature that exists around the Southern Women data concerns the order of presentation of the rows and columns. The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous. It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.  As \citet{freeman2003finding} notes, in the canonical original table included in \citet{davis1941}, both the row and column modes had been re-ordered in a presentation to reveal what DGG ascertained (using intuitive methods and in-depth knowledge of the case) was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.

To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns. In fact, as shown in Figure~\ref{fig:sw}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.\footnote{Using the DGG labeling, the chronologically ordered set of events are \{E11, E5, E2, E7, E12, E9, E3, E6, E10, E1, E14 E8, E4, E13\} \citep[p. 68]{everett2018measuring}.} To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}. This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.

Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns. Event ordering, however, matters for analyses that try to exploit the temporal properties of the data \citep[e.g.,][]{everett2018measuring, lerner2022dynamic}. Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to "W1" or "E8" in the other \citep[e.g.,][]{doreian1979evolution}. To forestall confusion, in what follows, I use the numeric labeling corresponding to the original DGG data table as reproduced in (among others) \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{fig:sw}. My presentation includes both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" (W11) rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.

\subsubsection{Reflective Score Trajectory}
Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections. The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis. As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to. However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections).

In the exhaustive methodological meta-analysis mentioned earlier, \citet{freeman2003finding} noted that ``the consensus of the analytic procedures is to assign women 1 through 9 to one group and women 10 through 18 to the other.'' Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover this same dominant partition reading from top to bottom. Figure~\ref{fig:g-refs} shows the corresponding bump chart for the events.  Clearly, $E14$ experiences the most dramatic improvement in status as we move to higher reflections. Relatively low ranked when it comes to the average number of memberships of its members, it increases in standing when considering (the average of) the average number of memberships of its members (and so forth). Notably, the equilibrium reflective centralities recover and ordering of groups columns that re-appears across many articles that have analyzed these data with events 1-6 separated from events 10-15 by events 7-9 \citep[see e.g.,][]{doreian2004generalized, kovacs2010generalized, lizardo2024two, everett2013dual}.

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
            \caption{Reflective centrality trajectories of persons.}
            \label{fig:p-refs}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/g-reflections.png}
            \caption{Reflective centrality trajectories of groups.}
            \label{fig:g-refs}
    \end{subfigure}
    \caption{Reflective centralities for persons and groups (even reflections)}
    \label{fig:refs}
\end{figure}

\section{CA and Dual Centrality in Two-Mode Networks} \label{sec:ca}
The reader may ask what the point of going through all these reflective scoring, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected. Recall that, as noted earlier in that paper, Bonacich also considered CA in passing but dismissed it as a tool for assigning scores to nodes in two-mode networks, noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}. While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network. In fact, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.

To see this, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:

\begin{equation}
    AC^B_p = \lambda C^B_g 
    \label{eq:bon1}
\end{equation}

\begin{equation}
    A^TC^B_g = \lambda C^B_p 
    \label{eq:bon2}
\end{equation}

Where $A$ is the original affiliation matrix. Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.

For people, these are given by:

\begin{equation}
    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
    \label{eq:faust1}
\end{equation}

And for groups:

\begin{equation}
    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
    \label{eq:faust2}
\end{equation}

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich dual-scoring approach, the centrality of people is equal to the sum of the centralities scores of the groups they join, and the centrality of groups is equivalent to the sum of the centralities of their members. This approach thus nicely captures the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode. Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members. We will return to this crucial point later.

\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
To see the connection between the reflective and Bonacich two-mode centralities more clearly, we can motivate the Bonacich two-mode centrality using the same ``reflective'' approach we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}. Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centrality for two-mode networks~\citep{bonacich1991simultaneous}, but it will help clarify the similarities and differences between the HH and the Bonacich approaches. Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:

\begin{equation}
   C^B_p(2) = \sum_g a_{pg}C^R_g(1)
   \label{eq:BR2_p}
\end{equation}

\begin{equation}
   C^B_g(2) = \sum_p a_{pg}C^R_p(1)
   \label{eq:BR2_g}
\end{equation}

Equation~\ref{eq:BR2_p} says that people are central when the sum of the number of members of the groups they belong to is large. Equation~\ref{eq:BR2_g} says groups are central when the sum of the number of memberships of the people who belong to them is large.

As before, we can keep on going and define a third-order reflection using the formulas:

\begin{equation}
   C^B_p(3) = \sum_g a_{pg}C^R_g(2)
   \label{eq:BR3_p}
\end{equation}

\begin{equation}
   C^B_g(3) = \sum_p a_{pg}C^R_p(2)
   \label{eq:BR3_g}
\end{equation}

Equation~\ref{eq:BR3_p} says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large. Equation~\ref{eq:BR3_g} says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large. Once again, we can keep going and define fourth order, fifth order, and higher reflections $C^R_p(4), C^R_p(4) \ldots C^R_p(q)$, where the centralities of nodes in one mode are based on the sums, of the sums, of the sums, of the centralities of nodes in the other mode.

More generally, and in parallel with equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}, the reflective Bonacich centralities for persons and groups are given by:

\begin{equation}
   C^B_p(q) = \sum_g a_{pg}C^R_g(q-1)
   \label{eq:BRq_p}
\end{equation}

\begin{equation}
   C^B_g(q) = \sum_p a_{pg}C^R_p(q-1)
   \label{eq:BRq_g}
\end{equation}

This says that the Bonacich reflective centrality at step $q$ is just the sum of the group centralities at step $q-1$ (for people) and the sum of the person centralities at step $q-1$ (for groups). Of course, these sums of sums would diverge to a larger and larger quantity at each step $q$. To prevent this and guarantee convergence, we can normalize the vector of reflective Bonacich centralities for persons and groups at each step $q > 1$ before calculating the subsequent sum at step $q+1$ as follows:

\begin{equation}
   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
   \label{eq:BRq_pn}
\end{equation}

\begin{equation}
   C^B_g(q) = \frac{C^B_g(q)}{||C^B_g(q)||_2}
   \label{eq:BRq_gn}
\end{equation}

Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}. Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations. These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach. Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power'' method of \citet{mises1929praktische}).

This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities. In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.\footnote{Note that this is precisely the difference between the Bonacich centrality in one-mode networks \citep{bonacich1972factoring} and the PageRank centrality \citep{brin1998anatomy}.}

\subsection{HH Reflections and Correspondence Analysis} \label{subsec:refeigen}
Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem. As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.

To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}. Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:

\begin{equation}
    \left(AA^T\right)C^B_p = \lambda C^B_p
    \label{eq:bon3}
\end{equation}

\begin{equation}
    \left(A^TA\right)C^B_g = \lambda C^B_g
    \label{eq:bon4}
\end{equation}

This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.

Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}. Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.

To see this, consider the $|P| \times |P|$ matrix $Dp$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell. In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell. Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).

For the people, we can do this by taking the original affiliation matrix and pre-multiplying it by $\mathbf{D}p^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each person $C^R_p(1)$ along the diagonals), yielding the row-stochastic matrix $\mathbf{P}_{pg}$ of dimensions $|P| \times |G|$:

\begin{equation}
    \mathbf{P}_{pg} = \mathbf{D}p^{-1}\mathbf{A}
    \label{eq:ca1}
\end{equation}

We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:

\begin{equation}
    \mathbf{P}_{gp} = \mathbf{D}_G^{-1}\mathbf{A}^T
    \label{eq:ca2}
\end{equation}

Both $\mathbf{P}_{pg}$ and $\mathbf{P}_{gp}$ are just centrality-weighted versions of the original affiliation matrix and its transpose; accordingly, nothing is stopping us from proceeding with the usual next step---due to \citet{breiger1974duality}---of calculating the degree-weighted \textit{projections} for people and groups from these matrices.

For the people, the degree-weighted projection is:

\begin{equation}
\mathbf{P}_{pp} = \mathbf{P}_{pg}\mathbf{P}_{gp}
    \label{eq:ca3}
\end{equation}

And for the groups:

\begin{equation}
    \mathbf{P}_{gg} = \mathbf{P}_{gp}\mathbf{P}_{pg}
    \label{eq:ca4}
\end{equation}

Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups. Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), but not symmetric, as can be seen in Figure~\ref{fig:dwproj}. Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person. Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix). The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/p-norm.png}
            \caption{Degree-weighted projection (People).}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/g-norm.png}
            \caption{Degree-weighted projection (Groups).}
            \label{fig:bon-reord}
    \end{subfigure}
    \caption{Degree-weighted projection matrices of the Southern Women data.}
    \label{fig:dwproj}
\end{figure}

To get to the main point of all of this, it can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as a solution of the eigensystem:

\begin{equation}
    \mathbf{P}_{pp}C^R_p = \lambda_2C^R_p
    %\left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
    \label{eq:dam1}
\end{equation}

\begin{equation}
    \mathbf{P}_{gg}C^R_g = \lambda_2C^R_g
    %\left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
    \label{eq:dam2}
\end{equation}\

With the $C^R$ scores obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the degree-weighted projection matrices. This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.  Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq. 9.17]{fouss2016algorithms}.\footnote{Note also that just like the eigenvector decomposition, the matrices implied by equations~\ref{eq:dam1} and~\ref{eq:dam2} have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix.}

More importantly, because $\mathbf{P}_{pp} = Dp^{-1}ADg^{-1}A^T$ and $\mathbf{P}_{gg} = Dg^{-1}A^TDp^{-1}A$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ and $\mathbf{P}_{pp}$ for people and $A^TA$ and $\mathbf{P}_{gg}$ for groups. The key difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.

This can be clearly seen if we express $\left(Dg^{-1}\mathbf{A}^TDp^{-1}\mathbf{A}\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:

\begin{equation}
    a_{pp'} = \sum_g\frac{a_{pg}a_{p'g}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_p}\sum_g\frac{a_{pg}a_{p'g}}{C^R_g(1)}
    \label{eq:mealy1}
\end{equation}

In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$. Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}. As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups). The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.

The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:

\begin{equation}
    a_{gg'} = \sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)C^R_g(1)} = 
    \frac{1}{C^R_g}\sum_p\frac{a_{pg}a_{pg'}}{C^R_p(1)}
    \label{eq:mealy2}
\end{equation}

Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}\mathbf{A}D_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}. As we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix for both the row and column objects.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/p-ca-ref-corr.png}
            \caption{CA/Reflection Correlation (Persons).}
            \label{fig:p-ca-ref}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/g-ca-ref-corr.png}
            \caption{CA/Reflection Correlation (Groups)}
            \label{fig:g-ca-ref}
    \end{subfigure}
    \caption{Scatter plot of the scores computed by the method of reflections ($q = 26$) on the y-axis and the first dimension of the Correspondence Analysis of the two-mode affiliation matrix on the x-axis, for both persons (a) and groups (b). }
    \label{fig:ca-ref-corr}
\end{figure}

Degree-pre-weighting does not radically alter the nature of the data. Indeed, it corresponds to moving from sums to averages, thus ``adjusting'' for the influence of person-activity and group size---a seemingly perennial issue in two-mode data analysis \citep[159ff]{bonacich1991simultaneous}. This can be seen by the fact that Equations~\ref{eq:mealy1} and~\ref{eq:mealy2} show that, substantively, what the degree pre-weighting does is that, for people, co-memberships count for more in determining interpersonal similarity when the group in question is small than when it is large. On the group side, shared members who do not have many affiliations count more in determining intergroup similarity than those with many affiliations.

Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical. In the Bonacich case, an eigendecomposition of the one-mode projection of the two-mode network; in the reflective case, an eigendecomposition of the degree-weighted one-mode projection of the same network. Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks. Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to. The (average) centrality of groups, in turn, is a function of their members' (average) centrality. The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
            \caption{Re-ordered Affiliation matrix (CA First Dimension).}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
            \caption{Re-ordered Affiliation matrix (Bonacich Centrality).}
            \label{fig:bon-reord`}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
            \caption{Re-ordered Affiliation matrix (Bonacich Second Dimension).}
            \label{fig:bon-reord}
    \end{subfigure}
    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
    \label{fig:ca-v-bon}
\end{figure}

\subsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ reflective scores reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).

The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side). In fact, as \citet[p. 206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.  The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

\subsection{Eigenvector Plot} \label{subsec:eigplot}
We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis. If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}. The Southern Women data clearly belong to the second category. In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other. Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position. Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach. This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/p-ca-eigvec.png}
            \caption{Eigenvector plot (Persons).}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/g-ca-eigvec.png}
            \caption{Eigenvector plot (Groups).}
            \label{fig:bon-reord}
    \end{subfigure}
    \caption{CA Eigenvector plot for persons (a) and groups (b). The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis. Insets show the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
    \label{fig:ca-eigvec}
\end{figure}

\subsection{Normalized Similarity Plot} \label{subsec:normsim}
To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one. Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups). As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$. This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition. These are shown as inset plots in Figure~\ref{fig:ca-eigvec} for both people and groups. Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
        \includegraphics[width=1.0\textwidth]{Plots/ca-corr-plot.png}
            \caption{CA correspondence plot.}
            \label{fig:ca-corr}
    \caption{CA similarity plots for persons (a) and groups (b) on the left and right top panel. The bottom panel (c) shows the correspondence plot for persons and groups with scores corresponding to the first dimension on the horizontal axis and scores corresponding to the second dimension on the vertical axis.}
    \label{fig:ca-sim}
\end{figure}

To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups). In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity. Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups. As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.

\subsection{CA Correspondence Plot} \label{subsec:caplot}
The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions. People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices. The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups. People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm). The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart. The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.

Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups). This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot. But this (still common) interpretation is off the mark. What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs. 2.13 and 2.14]{fouss2016algorithms}. As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups. In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
            \caption{Bonacich correspondence plot.}
            \label{fig:bon-corr}
    \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel. The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis.}
    \label{fig:bon-sim}
\end{figure}

\subsection{One-Mode Projection Matrix and Eigenvector Plot} \label{sec:eigplot}
This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$. But as we have seen, this is \textit{not} what CA is designed to do. Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes. Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).

To illustrate, Figures~\ref{fig:bon-sim}(a) and~\ref{fig:bon-sim}(b) shows the ``raw'' (unweighted by other-mode degree) similarity scores for persons ($a_{pp'} = \sum_g a_{pg}a_{p'g}$) and groups ($a_{gg'} = \sum_p a_{pg}a_{pg'}$), with the rows and columns sorted by the first eigenvector of the matrix, which is the usual Bonacich centrality score. Both similarity matrices reproduce the triangular, core-periphery structure we observed in the re-ordered affiliation matrix in Figure~\ref{fig:ca-v-bon}(b).  Figure~\ref{fig:bon-sim}(c) plots $C^B$ on the x-axis against the second eigenvector of the unweighted similarity matrix---a relatively unusual but not substantively unmotivated practice \citep{iacobucci2017eigenvector}.\footnote{Nodes are colored to a four-cluster k-means solution using the first six eigenvectors of the respective similarity matrices.}  We can see that the plot of the first two eigenvectors does a good job of recovering the raw connectivity structure of the Southern Women affiliation network, partitioning the core persons and groups (on the upper-right) from the more peripheral ones (on the lower-left).

Moreover, if all we are interested in is capturing a low-dimensional representation of which people have particular affinities for which events (regardless of people's activity levels or group sizes), then Figure~\ref{fig:bon-sim}(c) does a better job of that than the usual CA correspondence plot in Figure~\ref{fig:ca-sim}(c). For instance, $\left[Flora, Olivia\right]$ do have a special affinity for $\left[E11\right]$ and $\left[Nora, Katherine, Sylvia\right]$ for $\left[E10, E12\right]$. In the same way, core events like $\left[E8, E9, E10\right]$---shown on the lower half of the plot---are different from core events $\left[E3, E4, E5, E6\right]$---shown in the upper half. The former are more inclusive of peripheral members while the latter are more ``cliquish,'' including only core members.

This ``preferential attachment'' \citep{barabasi1999emergence} of core people to core events and peripheral people to peripheral events seems to be captured by the second dimension, left over after considering each node's Bonacich eigenvector centrality. For instance, $\left[E1, E2\right]$, although as poorly attended as other peripheral events, tends to include core members and thus appear closer to other core actors in the upper half of the plot. Similarly, the main difference between $\left[Ruth\right]$ and $\left[Sylvia\right]$, despite their comparable $C^B$ scores, is that the former's (closer to the upper half of the plot) event profile is composed mainly of core events. In contrast, peripheral events dominate the latter's attendance profile (shown in the bottom half of the plot), accordingly, $\left[Ruth\right]$ and $\left[Sylvia\right]$ are assigned to distinct clusters by the k-means algorithm.

\subsection{Correspondence Analysis and Generalized Relational Similarity} \label{subsec:cagensim}
As noted, there seems to be a relationship between the ordering of persons and groups produced by CA of a two-mode network and that produced by previous work using a ``generalized relational similarity'' (GRS) strategy. Recall that for objects (let us say people) to be similar according to the GRS criterion, they must have overlapping connections, and those links should go to objects in the other mode \textit{that are themselves similar}, where objects' similarities are given by their pattern of connections to objects in the other mode. This recursive definition of similarity thus recalls the classic distinction between structural and regular equivalence \citep{everett1994regular}. In the context of two-mode networks, \citet{kovacs2010generalized} proposed one such approach to defining a GRS for nodes in one and two-mode networks using a modified version of the correlation distance.\footnote{\citet{lizardo2024two} shows the connection between Kovacs's idea of generalized relations similarity and the two-mode network projection.}

An earlier effort to define a GRS for persons and groups in two-mode networks, one more relevant for a direct comparison with CA and the ``reflective'' approaches already considered, can be found in \citet{jeh2002simrank}. In that work, the authors dubbed their similarity measure ``SimRank.'' In the context of two-mode network analysis, the goal is to compute a matrix of similarities for each of the two-node sets, where the similarity of people is a function of the groups they belong to, and the similarity of groups is a function of the people who belong to them, making the similarity of persons and groups ``mutually reinforcing notions'' \citep[540]{jeh2002simrank}. Thus,

\begin{itemize}
    \item People are \textit{similar} if they belong to \textit{similar} groups.
    \item Groups are \textit{similar} if they share \textit{similar} members.
\end{itemize}

Which is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}. To accomplish this, \citet[540, eq. 2 and eq. 3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:

\begin{equation}
    S(p, p') = \frac{\alpha}{C^R(1)_pC^R(1)_{p'}}
    \sum_{i = 1}^{C^R(1)_p} \sum_{j = 1}^{C^R(1)_{p'}} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
    \label{eq:simrank1}
\end{equation}

Where everything is as before, and $g(i)_{i \in N(p)}$ is the $i^{th}$ group in the set of groups that person $p$ belongs to, $g(j)_{j \in N(p')}$ is the $j^{th}$ group in the set of groups that person $p'$ belongs to, and $\alpha$ is a free parameter obeying the restriction: $0 > \alpha < 1$. By construction, $S(p, p) = 1$, for all $p$. Thus, equation \ref{eq:simrank1} says that the SimRank similarity between two people is a function of the sum of the similarities of each unordered pair of groups they both belong to, weighted by the reciprocal of the products of their number of memberships multiplied by $\alpha$.

Likewise, for groups, the SimRank similarities are given by:

\begin{equation}
    S(g, g') = \frac{\alpha}{C^R(1)_gC^R(1)_{g'}}
    \sum_{i = 1}^{C^R(1)_g} \sum_{j = 1}^{C^R(1)_{g'}} 
    S\left(p(i)_{i \in N(g)}, p(j)_{j \in N(g')}\right)
    \label{eq:simrank2}
\end{equation}

Where $p(i)_{i \in N(g)}$ is the $i^{th}$ person in the set of members of group $g$, and $p(j)_{j \in N(g')}$ is the $j^{th}$ person in the set of members of group $g'$, and $S(g, g) = 1$. Thus, the SimRank similarity between two groups is a function of the sum of the similarities of each pair of people who belong to both groups, weighted by the reciprocal of the products of the number of members of each group multiplied by $\alpha$.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
        \includegraphics[width=0.9\textwidth]{Plots/simrank-v-ca.png}
            \label{fig:sim-v-ca}
    \caption{Simrank versus CA comparison. Panels (a) and (c) show the final SimRank similarity matrices for persons and groups (respectively) with rows and columns ordered according to the first CA dimension scores.  Panel (b) shows the Pearson correlation between the scores corresponding to the second eigenvector of the SimRank similarity matrix (on the y-axis) and the scores corresponding to the first CA dimension (on the x-axis) for persons, panel (d) shows the same correlation for groups.}
    \label{fig:simrank}
\end{figure}

In two-mode networks, SimRank scores for each pair of nodes across the two sets can be estimated via a simple algorithm, in which we first estimate $S(p, p')$ in equation \ref{eq:simrank1} using baseline values. Hence, only groups that two people share contribute to the initial values of $S(p, p')$ since only $S(g, g)>0$ at the outset. We then plug those values into equation \ref{eq:simrank2}, then loop back to equation \ref{eq:simrank1} with the resulting $S(g, g')$ values, and continue iterating until convergence---generally achieved after five iterations \citep{jeh2002simrank}, which is confirmed here for the Southern Women data.

Note that equations~\ref{eq:simrank1} and~\ref{eq:simrank2} share a formal similarity with HH's ``method of reflections'' discussed earlier, in particular, the fact that both compute quantities based on nodes in the other mode averaged by the degree of nodes in the focal mode. The key difference is that SimRank works directly with pairwise comparisons between node dyads \citep{jeh2002simrank}. Nevertheless, this double weighting by degree should make us suspect that the results of the SimRank similarity analysis would not be too far from those obtained via CA, given the mathematical equivalence of CA and the method of reflections \citep{mealy2019interpreting, van2021correspondence}.

And indeed, they are not. Figure~\ref{fig:simrank}(a) shows the original affiliation matrix, this time with rows and columns re-arranged according to the values of the main non-trivial eigenvector (corresponding to the second-largest eigenvalue) of the SimRank similarity matrices---with $\alpha$ set to $0.8$---for the row and column nodes. Like in Figure~\ref{fig:ca-v-bon}(a), this reshuffling reveals the block-diagonal structure separating the two communities of persons and groups revealed by CA, suggesting that CA and SimRank uncover a similar underlying partitioning of the two-mode network's community structure.

Figure~\ref{fig:simrank}(b) and Figure~\ref{fig:simrank}(d) show a plot similar to that shown in Figures~\ref{fig:ca-eigvec}(a) and Figure~\ref{fig:ca-eigvec}(b) but this time with the main eigenvector of SimRank similarity matrices of both persons and groups on the x-axis and the corresponding rank on the y-axis. Looking at the plots from left to right, we can see that the partitioning of the node sets on the first informative eigenvector is substantively equivalent to those revealed by the first CA eigenvector, separating similar blocks of people and events.

Figure~\ref{fig:simrank}(f) shows the correspondence plot of the first and second eigenvectors of the SimRank similarity matrices for both persons and groups. Once again, this plot is substantively equivalent (in terms of the internode groupings and distances) to that shown in Figure~\ref{fig:ca-sim}(c), suggesting that the first-CA dimensions of the two-mode network recover clusters of ``similar'' entities in the two node sets, where internode similarity is a generalized relational similarity as defined earlier. Thus, [$Flora, Olivia$] appear close because they choose similar groups, as do [$Sylvia, Katherine, Helen$]. In the same way, groups [$E9, E11$] are close in the plot because they are chosen by similar people, as are events [$E10, E12, E13, E14$]. Figures~\ref{fig:simrank}(g) and (h) show a regression plot of the first CA eigenvector for persons (top) and groups (bottom) on the x-axis and the first eigenvector of the SimRank similarity matrix for both persons and groups on the y-axis. As we can see, the CA and SimRank ordering of nodes along the first dimension agree quite closely ($r = 0.98$ for persons and $r = 0.99$ for groups), confirming that these first eigenvectors capture the same (GSR) information across the two approaches.

\section{Discussion and Concluding Remarks} \label{sec:disc}
This paper reconsiders the role of CA in the analysis of two-mode network data. We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix. Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}. I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums. This exercise clarifies the links between CA and the dual two-mode eigenvector centrality more coherently and systematically, an issue that was left open and somewhat unclear in Bonacich's \citeyearpar{bonacich1991simultaneous} classic paper. Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.

But CA does more than reveal a latent dual ordering of nodes in the two-mode network. In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode. This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix. When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data. This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events \citep{everett2013dual}.

This is another way in which CA and the Bonacich dual centrality approach systematically differ; one is geared to uncovering blocks of actors with similar connections to events in the other mode (and vice versa), while the other reveals blocks of actors who are the most active and who attend the largest events. Both are, of course, legitimate ways of analyzing the structure of a two-mode network. Still, they differ in terms of the structural patterns that they are sensitive to, with the CA approach closer to a community partition where nodes that are surprisingly similar end up in the same clusters---where ``surprising'' means similarity based on deviations from a suitable null model based on random mixing given their first-order degrees.

But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to? Here, we show that CA recovers a type of generalized relational similarity (GRS). That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar. Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach. Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.

Overall, the preceding has shown that CA can be upgraded from a method designed to generate joint plots and visualization of two-mode data to one that can be seen as more ``central'' to the usual social-network-analytic tasks, like ordering the nodes in the two sets according to some substantively meaningful rank---centrality analysis---or finding sets of similar actors in the network (subgroup or community detection). At the very least, it seems like the Bonacich style dual centrality based on the eigenvector decomposition of the raw affiliation matrix should not be the default ``reflective'' centrality approach unless the analyst has the explicit analytic goal of exploring center-periphery partitioning in the network.

A better approach, similar to the one exemplified here, would be to present a comparison of the reflective centralities obtained by CA and Bonacich side-by-side to see whether the underlying display a substantively interesting similarity partitioning in addition to any core-periphery ordering. Of course, suppose the analyst is more interested in such a ``subgroup'' analytic approach. In that case, the Bonacich dual centrality approach is less helpful (even if multiple dimensions are considered), and CA should be the first line of attack.

%% Loading bibliography style file]
%\bibliographystyle{model1-num-names}
\bibliographystyle{cas-model2-names}
% Loading bibliography database
\bibliography{ca.bib}
\end{document}

The Elsevier cas-dc class is based on the
standard article class and supports almost all of the functionality of
that class. In addition, it features commands and options to format the
\begin{itemize} \item document style \item baselineskip \item front
matter \item keywords and MSC codes \item theorems, definitions and
proofs \item labels of enumerations \item citation style and labeling.
\end{itemize}

This class depends on the following packages
for its proper functioning:

\begin{enumerate}
\itemsep=0pt
\item {natbib.sty} for citation processing;
\item {geometry.sty} for margin settings;
\item {fleqn.clo} for left aligned equations;
\item {graphicx.sty} for graphics inclusion;
\item {hyperref.sty} optional packages if hyperlinking is
  required in the document;
\end{enumerate}

All the above packages are part of any
standard \LaTeX{} installation.
Therefore, the users need not be
bothered about downloading any extra packages.

\section{Installation}

The package is available at author resources page at Elsevier
(\url{http://www.elsevier.com/locate/latex}).
The class may be moved or copied to a place, usually,\linebreak
\verb+$TEXMF/tex/latex/elsevier/+, %$%%%%%%%%%%%%%%%%%%%%%%%%%%%%
or a folder which will be read                   
by \LaTeX{} during document compilation.  The \TeX{} file
database needs updation after moving/copying class file.  Usually,
we use commands like \verb+mktexlsr+ or \verb+texhash+ depending
upon the distribution and operating system.

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. 
You are recommended to conform your choice to the journal you 
are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use Bib\TeX\ to generate your bibliography and include DOIs
whenever available.

Here are two sample references: 
\cite{Fortunato2010}
\cite{Fortunato2010,NewmanGirvan2004}
\cite{Fortunato2010,Vehlowetal2013}

\section{Floats}
{Figures} may be included using the command,\linebreak 
\verb+\includegraphics+ in
combination with or without its several options to further control
graphic. \verb+\includegraphics+ is provided by {graphic[s,x].sty}
which is part of any standard \LaTeX{} distribution.
{graphicx.sty} is loaded by default. \LaTeX{} accepts figures in
the postscript format while pdf\LaTeX{} accepts {*.pdf},
{*.mps} (metapost), {*.jpg} and {*.png} formats. 
pdf\LaTeX{} does not accept graphic files in the postscript format.

\begin{figure}
	\centering
		\includegraphics[scale=.75]{figs/Fig1.pdf}
	\caption{The evanescent light - $1S$ quadrupole coupling
	($g_{1,l}$) scaled to the bulk exciton-photon coupling
	($g_{1,2}$). The size parameter $kr_{0}$ is denoted as $x$ and
	the \PMS is placed directly on the cuprous oxide sample ($\delta
	r=0$, See also Table \protect\ref{tbl1}).}
	\label{FIG:1}
\end{figure}

The \verb+table+ environment is handy for marking up tabular
material. If users want to use {multirow.sty},
{array.sty}, etc., to fine control/enhance the tables, they
are welcome to load any package of their choice and
{cas-dc.cls} will work in combination with all loaded
packages.

\begin{table}[width=.9\linewidth,cols=4,pos=h]
\caption{This is a test caption. This is a test caption. This is a test
caption. This is a test caption.}\label{tbl1}
\begin{tabular*}{\tblwidth}{@{} LLLL@{} }
\toprule
Col 1 & Col 2 & Col 3 & Col4\\
\midrule
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
12345 & 12345 & 123 & 12345 \\
\bottomrule
\end{tabular*}
\end{table}

\section[Theorem and ...]{Theorem and theorem like environments}

{cas-dc.cls} provides a few shortcuts to format theorems and
theorem-like environments with ease. In all commands the options that
are used with the \verb+\newtheorem+ command will work exactly in the same
manner. {cas-dc.cls} provides three commands to format theorem or
theorem-like environments:

\begin{verbatim}
 \newtheorem{theorem}{Theorem}
 \newtheorem{lemma}[theorem]{Lemma}
 \newdefinition{rmk}{Remark}
 \newproof{pf}{Proof}
 \newproof{pot}{Proof of Theorem \ref{thm2}}
\end{verbatim}

The \verb+\newtheorem+ command formats a
theorem in \LaTeX's default style with italicized font, bold font
for theorem heading and theorem number at the right hand side of the
theorem heading.  It also optionally accepts an argument which
will be printed as an extra heading in parentheses.

\begin{verbatim}
  \begin{theorem} 
   For system (8), consensus can be achieved with 
   $\|T_{\omega z}$ ...
     \begin{eqnarray}\label{10}
     ....
     \end{eqnarray}
  \end{theorem}
\end{verbatim}

\newtheorem{theorem}{Theorem}

\begin{theorem}
For system (8), consensus can be achieved with 
$\|T_{\omega z}$ ...
\begin{eqnarray}\label{10}
....
\end{eqnarray}
\end{theorem}

The \verb+\newdefinition+ command is the same in
all respects as its \verb+\newtheorem+ counterpart except that
the font shape is roman instead of italic.  Both
\verb+\newdefinition+ and \verb+\newtheorem+ commands
automatically define counters for the environments defined.

The \verb+\newproof+ command defines proof environments with
upright font shape.  No counters are defined.

\section[Enumerated ...]{Enumerated and Itemized Lists}
{cas-dc.cls} provides an extended list processing macros
which makes the usage a bit more user friendly than the default
\LaTeX{} list macros.   With an optional argument to the
\verb+\begin{enumerate}+ command, you can change the list counter
type and its attributes.

\begin{verbatim}
 \begin{enumerate}[1.]
 \item The enumerate environment starts with an optional
   argument `1.', so that the item counter will be suffixed
   by a period.
 \item You can use `a)' for alphabetical counter and '(i)' 
  for roman counter.
  \begin{enumerate}[a)]
    \item Another level of list with alphabetical counter.
    \item One more item before we start another.
    \item One more item before we start another.
    \item One more item before we start another.
    \item One more item before we start another.
\end{verbatim}

Further, the enhanced list environment allows one to prefix a
string like `step' to all the item numbers.

\begin{verbatim}
 \begin{enumerate}[Step 1.]
  \item This is the first step of the example list.
  \item Obviously this is the second step.
  \item The final step to wind up this example.
 \end{enumerate}
\end{verbatim}

\section{Cross-references}
In electronic publications, articles may be internally
hyperlinked. Hyperlinks are generated from proper
cross-references in the article.  For example, the words
\textcolor{black!80}{Fig.~1} will never be more than simple text,
whereas the proper cross-reference \verb+\ref{tiger}+ may be
turned into a hyperlink to the figure itself:
\textcolor{blue}{Fig.~1}.  In the same way,
the words \textcolor{blue}{Ref.~[1]} will fail to turn into a
hyperlink; the proper cross-reference is \verb+\cite{Knuth96}+.
Cross-referencing is possible in \LaTeX{} for sections,
subsections, formulae, figures, tables, and literature
references.

\section{Bibliography}

Two bibliographic style files (\verb+*.bst+) are provided ---
{model1-num-names.bst} and {model2-names.bst} --- the first one can be
used for the numbered scheme. This can also be used for the numbered
with new options of {natbib.sty}. The second one is for the author year
scheme. When  you use model2-names.bst, the citation commands will be
like \verb+\citep+,  \verb+\citet+, \verb+\citealt+ etc. However when
you use model1-num-names.bst, you may use only \verb+\cite+ command.

\verb+thebibliography+ environment.  Each reference is a\linebreak
\verb+\bibitem+ and each \verb+\bibitem+ is identified by a label,
by which it can be cited in the text:

\noindent In connection with cross-referencing and
possible future hyperlinking it is not a good idea to collect
more that one literature item in one \verb+\bibitem+.  The
so-called Harvard or author-year style of referencing is enabled
by the \LaTeX{} package {natbib}. With this package the
literature can be cited as follows:

\begin{enumerate}[\textbullet]
\item Parenthetical: \verb+\citep{WB96}+ produces (Wettig \& Brown, 1996).
\item Textual: \verb+\citet{ESG96}+ produces Elson et al. (1996).
\item An affix and part of a reference:\break
\verb+\citep[e.g.][Ch. 2]{Gea97}+ produces (e.g. Governato et
al., 1997, Ch. 2).
\end{enumerate}

In the numbered scheme of citation, \verb+\cite{<label>}+ is used,
since \verb+\citep+ or \verb+\citet+ has no relevance in the numbered
scheme.  {natbib} package is loaded by {cas-dc} with
\verb+numbers+ as default option.  You can change this to author-year
or harvard scheme by adding option \verb+authoryear+ in the class
loading command.  If you want to use more options of the {natbib}
package, you can do so with the \verb+\biboptions+ command.  For
details of various options of the {natbib} package, please take a
look at the {natbib} documentation, which is part of any standard
\LaTeX{} installation.

\appendix
\section{My Appendix}
Appendix sections are coded under \verb+\appendix+.

\verb+\printcredits+ command is used after appendix sections to list 
author credit taxonomy contribution roles tagged using \verb+\credit+ 
in frontmatter.

\printcredits

%\vskip3pt

\bio{}
Author biography without author photo.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
\endbio

\bio{figs/pic1}
Author biography with author photo.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
\endbio

\bio{figs/pic1}
Author biography with author photo.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
Author biography. Author biography. Author biography.
\endbio